Statistics for complex graphs: an algorithmic approach

### 2014-10-22 Deliverable 13Q2/src/13q2.tex

%\usepackage{amsthm}
\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}

@@

We have scaled up the results from the previous work package in data volume and added various complexity measures. We will process more datasets, and attempt to establish interesting empirical results based on compression and complexity analysis. The existing implementations of complexity measures from 2012Q3 are included as a comparison.

The purpose of this deliverable is threefold. It validates that our methods, such as they are, can be used to provide useful empirical results. It will serve as a test of our software and make sure that it is maintained (preventing 'code rust'). And finally, it will provide us with feedback for which features of our code are interesting enough to scale up in Q3.

\section{Datasets}

\section{Complexity measures}

\section{Results}

\section{Interpretation}

>>>>

%\usepackage{amsthm}
\usepackage{charter}
\usepackage{eulervm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{hyperref}
\usepackage{lscape}

\usepackage{authordate1-4}

@@

\maketitle

We have scaled up the results from the previous deliverable in data volume and added various complexity measures. We will process more datasets, and attempt to establish interesting empirical results based on compression and complexity analysis. The existing implementations of complexity measures from 2012Q3 are included as a comparison.

The purpose of this deliverable is threefold. It validates that our methods, such as they are, can be used to provide useful empirical results. It will serve as a test of our software and make sure that it is maintained (preventing `code rust'). Finally, it will provide us with feedback for which features of our code are interesting enough to scale up in quarter 3.

Parts of this document have been taken and adapted from the deliverables of 2012 quarter 2 and quarter 3.

\section*{Code}

The code is written in Java. The JUNG graph library has been replaced with an alternative implemented for the project. All code is available from GitHub in the following repsitory. It is released under an MIT open source license. 
\begin{description}
\item[Lilian] (\url{https://github.com/pbloem/Lilian}) Implementation of the complexity measures, including some tools for running experiments. This repository also contains a great deal of code unrelated to this deliverable. The relevant code resides mostly in the package \texttt{org.lilian.graphs}.
\end{description}

\section{Data}

We test these complexity measures on the following datasets:

\subsection*{Technological networks}

\subsection{An Erdos-Renyi random graph}

A random graph consisting of 1000 nodes, with connection probability 0.5.

\subsection{A preferential attachment random graph}

A random graph consisting of 1000 nodes, generated by the Barabasi Albert algorithm, with 2 connection added per iteration.

\subsubsection*{The internet}

The internet is a classic example of a large, complex graph. So large, in fact that we have little hope of seeing a full snapshot of the whole network at once. The best we can do is to capture some part of it using traceroutes. We use a dataset downloaded from the personal page of Mark Newman (\url{http://www-personal.umich.edu/~mejn/netdata/}), which contains around 23 thousand nodes.

\subsubsection*{Commit social web}

This dataset was produced by the Data2Semantics project to map out the relations within the commit project.

\subsubsection*{IAFB Dataset}

This dataset was included to test our methods on RDF data. We take the dataset from \cite{bloehdorn2007kernel} and disregard any predicates that are not part of the domain specific ontology (http://swrc.ontoware.org/ontology). Named resources, blank nodes and literals are translated to vertices and predicates are translated to edges.

\subsubsection*{The world wide web}

The world wide web, while closely associated with the internet, is a fundamentally different type of network. For example, a link between two systems on the internet will bear some relation to physical proximity of the systems. Linking to a website that is hosted far away costs nothing.

The WWW dataset is know as www-barabasi, and is a common dataset in graph analysis.

\subsection*{Social networks}
\subsubsection*{Epinions}

Epinions.org is a review site where users can choose to `trust' one another. these trust relations are then used to determine the value of a given review. The total network, downloaded form \url{http://snap.stanford.edu/data/soc-Epinions1.html} contains about 76 thousand nodes.

Originally used in \cite{richardson2003trust}.

\subsubsection*{Pokec relationship network}

Pokec is a Slovakian social networking website. This dataset describes the social relations on the site.

Downloaded from \url{http://snap.stanford.edu/data/soc-pokec.html}. Citations: \cite{takacdata}

\section*{Experiments}

\subsubsection*{Basic complexity measures}

To show the scale of dataset that we can analyse, we perform some basic graph analyses on an datasets of increasing size.

\begin{tabular}{|l | r r r r r r r r |}
  \hline                       
  & $n$ & $l$ & $d_\mu$ & $d_\sigma$ & $W$ & $\omega$ & $\alpha$ & $t$ \\ 
  \hline
er  & $10000$ & $50045$ & $10.009$ & $3.14$ & $10000$ & $0.68$ & $0.0$ & $13.7 s$ \\ 
pa  & $10000$ & $19997$ & $4.0$ & $6.09$ & $10000$ & $0.145$ & $-0.06$ & $11.2 s$ \\ 
internet  & $22963$ & $48436$ & $4.29$ & $32.94$ & $22963$ & $0.030$ & $-0.20$ & $28.7 s$ \\ 
commit  & $18117$ & $56031$ & $6.19$ & $74.83$ & $18117$ & $0.065$ & $-0.13$ & $43.4 s$ \\ 
iafb  & $8318$ & $61255$ & $14.73$ & $69.87$ & $8318$ & $0.13$ & $-0.26$ & $20.6 s$ \\ 
epinions  & $75877$ & $508837$ & $13.41$ & $52.66$ & $75877$ & $0.11$ & $-0.011$ & $15.6 s$ \\ 
barabasi  & $325729$ & $1497135$ & $9.19$ & $48.38$ & $325729$ & $0.050$ & $-0.043$ & $30.0 s$ \\ 
skitter  & $1696415$ & $11095298$ & $13.08$ & $136.86$ & $1694616$ & $0.11$ & $-0.081$ & $7m 36 s$ \\ 
patents  & $6009555$ & $16518948$ & $5.5$ & $9.33$ & $3764117$ & $0.23$ & $0.17$ & $13h 27m 27s$ \\ 
social  & $1632804$ & $30622564$ & $37.51$ & $59.52$ & $1632803$ & $0.49$ & $ 5.3628-4 $ & $33m 34s$ \\ 
  \hline  
\end{tabular}

With the following meanings:
\begin{description}
\item[$n$] The number of nodes in the graph.
\item[$l$] The number of links in the graph.
\item[$d_\mu$] The mean degree.
\item[$d_\sigma$] The standard deviation of the degree distribution.
\item[$W$] The size of the largest weakly connected component.
\item[$\omega$] The wing width ratio \cite{kang2011beyond}.
\item[$\alpha$] The Assortativity
\item[$t$] The total running time of all analyses run for this report. \footnote{note that for graphs with more than 50000 nodes, the AM compressor wasn't run).}
\end{description}

\subsubsection*{Slash-and-burn and spyplots}

We implemented and studied the Slash-and-burn algorithm \cite{kang2011beyond}, and the spyplot method of visualization. This algorithm calculates an order of nodes that is likely to benefit compression (and computation) on real-world graphs.

The following plots show adjacency matrices (spyplots) for the datasets mentioned above.

\begin{table}[h]
\begin{tabular}{l | c c c c}
\hline
 & random & given & degree & sb \\
\hline
er
&
    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/er/adjacency-matrix-slashburn-ordering.png} \\

\hline
pa
&
    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/pa/adjacency-matrix-slashburn-ordering.png} \\

\hline
internet
&
    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/internet/adjacency-matrix-slashburn-ordering.png} \\

\hline
commit
&
    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/commit/adjacency-matrix-slashburn-ordering.png} \\

\hline
aifb
&
    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/aifb/adjacency-matrix-slashburn-ordering.png} \\

\hline
epinions
&
    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/epinions/adjacency-matrix-slashburn-ordering.png} \\

\hline
barabasi
&
    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/barabasi/adjacency-matrix-slashburn-ordering.png} \\
    \hline
\end{tabular}
\end{table}

\begin{table}[h]
\begin{tabular}{l | c c c c}
\hline
 & random & given & degree & sb \\
\hline
skitter
&
    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/skitter/adjacency-matrix-slashburn-ordering.png} \\

\hline
patents
&
    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/patents/adjacency-matrix-slashburn-ordering.png} \\

\hline
pokec
&
    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-random-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-given-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-degree-ordering.png}
& 
    \includegraphics[width=0.23\textwidth]{../img/pokec/adjacency-matrix-slashburn-ordering.png} \\

\hline

\end{tabular}
\end{table}

\subsubsection*{Compression tests}

Our ultimate goal is to study compression of graphs as a means of performing inference and complexity-analysis.\footnote{See also \cite{bloemcompression}, which was submitted as a deliverable this quarter.} To this end, we study the basic methods of representing graphs in a bitstring, and the effect that node ordsering has on the performance. We test the following methods
\begin{description}
\item[edge list (el)] The graph is stored as a list of prefix-coded pairs of integers representing the graph's edges. 
\item[neighbor list (nl)] The graph is stored as a list of lists of indices. Each list represents the neighbors of a given node. The indices are incremented by one so that the index 0 can be used as a delimiter.
\item[adjacency matrix (am)] We generate an adjacency matrix of $\frac{n^2 + n}{2}$ cells for an undirected graph or $n^2$ cells for a directed graph and flatten this into a bitstring. We zip the resulting string to exploit simple structures within it. We add a single bit to distinguish between directed and underected graphs so that the code is self-delimiting.
\end{description}

\begin{landscape}
\hspace{-1cm}
\begin{tabular}{|l | r r r | r r r | r r r | r r r |}
\hline
\multicolumn{1}{|c}{} & \multicolumn{3}{c}{natural} & \multicolumn{3}{c}{random} & \multicolumn{3}{c}{degree} & \multicolumn{3}{c|}{slashburn} \\
  \hline                       
  & el & nl & am & el & nl & am & el & nl & am & el & nl & am \\ 
  \hline
 er & 1538162 & 816534 & 146338 & 1537724 & 816010 & 146685 & 1494122 & 800123 & 144941 & 1495829 & 801555 & 143144 \\ 
 pa & 558505 & 317266 & 63798 & 615444 & 332357 & 68439 & 546885 & 318000 & 61711 & 551301 & 318136 & 58245 \\
 internet & 1329380 & 808430 & & 1594604 & 860025 & 165759 & 1190776 & 777404 & 114297 & 1224515 & 798488 &  \\
 commit & 1609616 & 673296 &  & 1817950 & 876699 & 215356 & 1370638 & 556954 & 135640 & 1411396 & 558630 &  \\
 aifb & 1583268 & 745875 & 70518 & 1844193 & 927728 & 155225 & 1366585 & 591342 & 69044 & 1403070 & 595136 & 69086 \\
 epinions & 1.5760E7 & 8091701 &  & 1.9181E7 & 9670775 &  & 1.4272E7 & 7530653 &  & 1.4720E7 & 7642574 &  \\
 barabasi & 6.3445E7 & 3.2238E7 & & 6.3769E7 & 3.2175E7 & & 5.0114E7 & 2.5458E7 & & 5.5466E7 & 2.83587E7 & \\
 skitter & 4.84613E8 & 2.2962E8 & & 5.2691E8 & 2.6509E8 &  & 4.2595E8 & 1.8885E8 & & 4.3375E8 & 1.9100E8 &  \\
 patents &  8.8071E8 &  4.5034E8  & & 8.4574E8 & 4.2886E8 &  & 7.6607E8 & 3.8617E8 &  & 7.6810E8 & 3.8725E8 &  \\
 pokec &  1.3859E9 &  6.9489E8  &  & 1.4513E9 & 7.2733E8 & & 1.3180E9 & 6.6017E8 &  & 1.3200E9 & 6.6129E8 &  \\
  \hline  
\end{tabular}
\end{landscape}
\section{Acknowledgements}
This research was supported by the Dutch national program COMMIT.

\bibliographystyle{authordate3}
\bibliography{citations}

### 2014-10-22 Nickname problem/src/nickname.tex

\title{The Nickname Problem}
\date{\today}

% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

@@

Note that this is a natural property of basic computational metaphors. We can always take part of the argument of one function and hardcode it into the program at the cost of its length, and some additional bits to add it to the argument before processing.

@@

\begin{lma}
Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions. For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq c(\max(i^*, j^*)) + O(1)$. Where $c(x) \leq 2 \log(|x|) + 1$ is the overhead of prefix encoding $x$
\end{lma}
\begin{proof}
Let $\phi_u$ be a universal function for $\psi$, such that $\phi_u(\bar{\imath}x) \simeq \psi_i(x)$. Since $\psi$ is computable this function exists in $\phi$. We have $\psi_{j^*}(x) \simeq \phi_u(\overline{\jmath^*}x)$. By the compactness of $\phi$, we have some $h$ with $|h| \leq |u| + |\overline{\jmath^*}| + O(1)$ such that $\phi_{h}(x) \simeq \phi_u(\overline{\jmath^*}x) \simeq \psi_{j^*}(x)$.

Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1) = |j*| + c(\jmath^*) + O(1)$. Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |i^*| + c(i^*) + O(1)$. Combining these, we get the desired result.
\end{proof}

This lemma tells us that so long as we restrict our numberings to be compact and computable, we can never gain more than a logarithmic difference between model sizes. In the limit as $i \rightarrow \infty$, this term represents a vanishing proportion.

>>>>

\title{The Nickname Problem}
\date{\ldots}
% non-indented, spaced paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

@@

Note that this is a natural property of basic computational metaphors. We can always take part of the argument of one function and hardcode it into the program at the cost of its prefix-coded length, and some additional bits to add it to the argument before processing.

@@

\begin{lma}
Let $\{\phi_i\}$ and $\{\psi_j\}$ be two compact, computable enumerations of the computable functions. For a given computable function $\phi_i \simeq \psi_j$,  we have $\mbox{abs}(|i^*| - |j^*|) \leq \max(|\overline{\imath^*}|, |\overline{\jmath^*}|) + O(1)$. That is, the difference between the sizes of the minimal models is in $O(\log(\max(i, j)))$
\end{lma}
\begin{proof}
Let $\phi_u$ be a universal function for $\psi$, such that $\phi_u(\bar{\imath}x) \simeq \psi_i(x)$. Since $\psi$ is computable this function exists in $\phi$. We have $\psi_{j^*}(x) \simeq \phi_u(\overline{\jmath^*}x)$. By the compactness of $\phi$, we have some $h$ with $|h| \leq |u| + |\overline{\jmath^*}| + O(1)$ such that $\phi_{h}(x) \simeq \phi_u(\overline{\jmath^*}x) \simeq \psi_{j^*}(x)$.

Thus $|i^*| \leq |h| = |u| + |\overline{\jmath^*}| + O(1)$. Reversing $\phi$ and $\psi$ gives us $|j^*| \leq |h'| = |u'| + |\overline{\imath^*}| + O(1)$. Combining these, we get the desired result.
\end{proof}

This lemma tells us that so long as we restrict our indices to compact and effective, we can never gain more than a logarithmic difference between model sizes. In the limit as $i \rightarrow \infty$, this term represents a vanishing proportion.

### 2014-10-22 Overview/src/overview.tex

\title{The question of interesting information}
\date{\today}

@@

\begin{abstract}
\noindent Ever since the introduction of absolute measures of information content---such as Information Entropy and Kolmogorov Complexity---there have been  criticisms of the idea that these measures capture the informal notion of  information. The argument is that the signals which, under these measures, contain the highest amount of information are by definition completely unpredictable, and therefore cannot contain any interesting information. There is an informal notion of information content, which is not captured by these absolutes, a kind of useful or meaningful information content. We review the literature available on this topic and provide a distillation of common themes and possible connections.
\end{abstract}

@@

Consider the following, illustrative example. An old-fashioned, black-and-white television set is turned off. The screen is black. This signal represents a complete lack of information. It is highly predictable. At every point we can predict exactly the next state of the signal. Describing this signal would be very simple.

@@

\begin{quotation}
A fundamentalproblem for statistical mechanics is to explain why dissipative systems (those in which entropy is continually being produced and removed to the surroundings) tend to undergo ``self-organisation,'' a spontaneous increase of structural complexity, of which the most extreme example is the origin and evolution of life. The converse principle, namely that nothing very interesting is likely to happen in a system in thermal equilibrium, is reflected in the term ``heat death.'' In the modern world view, thermodynamic driving forces, such as the temperature difference between the hot sun and the cold night sky, have taken over one of the functions of God: they make matter transcend its clod-like nature andbehave instead in dramatic and unforeseen ways, for example molding itself into thunderstorms, people, and umbrellas.
\end{quotation}

@@

Complexity, order, etc.
\section*{Preliminaries}

\subsection*{Information Entropy}

\subsection*{Kolmogorov Complexity}

\subsection*{Two-part coding}

\subsection*{Dynamical systems}

@@

\begin{itemize}
  \item Level of formalization
  \item The objectivity of interestingness
  \item Broadness. What types of data are covered, and what models and patterns are covered? 
  \item Context-dependence. Context plays an undeniable part in any notion of beauty or interestingness. Are contextual considerations acknowledged and given a place in the framework?
\end{itemize}

>>>>

\title{The question of interestingness}
\date{\today}

@@

\begin{abstract}
\noindent Ever since the introduction of absolute measures of information content---such as Information Entropy and Kolmogorov Complexity---there have been  criticisms of the idea that these measures capture the informal notion of  information. The argument is that the signals which, under these measures, contain the highest amount of information are of absolutely no use or interest to any observer. There is an informal notion of information content, which is not captured by these absolutes, a kind of useful or meaningful information content. We review the literature available on this topic and provide a distillation of common themes and possible connections.
\end{abstract}

@@

Consider the following example. An old-fashioned, black-and-white television set is turned off. The screen is black. This signal represents a complete lack of information. It is highly predictable. Describing this signal would be very simple.

@@

\begin{quotation}
A fundamental problem for statistical mechanics is to explain why dissipative systems (those in which entropy is continually being produced and removed to the surroundings) tend to undergo ``self-organisation,'' a spontaneous increase of structural complexity, of which the most extreme example is the origin and evolution of life. The converse principle, namely that nothing very interesting is likely to happen in a system in thermal equilibrium, is reflected in the term ``heat death.'' In the modern world view, thermodynamic driving forces, such as the temperature difference between the hot sun and the cold night sky, have taken over one of the functions of God: they make matter transcend its clod-like nature andbehave instead in dramatic and unforeseen ways, for example molding itself into thunderstorms, people, and umbrellas.
\end{quotation}

\section{Preliminaries: The question of complexity}

This section discusses the matter of what makes data complex. Most approaches to what makes data interesting build on these concepts. We review them here briefly, to briefly discuss our preliminaries and to define a uniform notation for the rest of the paper.

\subsection{Entropy}

In physical systems; in probability and coding.

\subsection{Kolmogorov Complexity}

Conditional complexity, Resource-bounded, MDL, two-part coding.

\section{Approaches to Interestingness}

@@

Complexity, order, etc.

@@

What questions should we ask of a proposed measure?
\begin{itemize}
  \item Level of formalization
  \item The objectivity of interestingness
  \item Broadness. What types of data are covered, and what models and patterns are covered? 
  \item Context-dependence. Context plays an undeniable part in any notion of beauty or interestingness. Are contextual considerations acknowledged and given a place in the framework?
\end{itemize}

checklist per paper:
\begin{itemize}
  \item Check google scholar for citations to the paper.
  \item Walk through the citations in the paper.
\end{itemize}

### 2014-10-22 PhD Thesis/src/thesis.tex

\chapter{Introduction}

@@

\section{The complex graph}

% Complex graphs are not just funny things to look at, the are the fundamental unit of science in the 21st century.

@@

\begin{summary}We review the literature on complex graphs, and the approaches that have been used so far to provide statistical analyses on them.
\end{summary}

% Algorithmic statistics is how you stud graphs. Nothing else makes sense.

\chapter{Algorithmic Statistics: Bridging the gap between Kolmogorov complexity and MDL}

\begin{summary}MDL and Kolmogorov complexity are two offshoots of the same basic idea: that good compression makes good statistics. MDL takes this as an axiom and builds on it, while Kolmogorov complexity makes only the assumption that the data is produced by a computational process. The two methods are not incompatible, but they have produced very different results and techniques. In order to have access to the best of both worlds, we start with the rigorous theoretical foundations of Kolmogorov complexity and build a platform that will allow us to fold in the techniques of MDL. We will revisit the principle of two-part coding, considered to be a usable, but cumbersome technique in MDL and show that when combined with Kolmogorov complexity it has some deep advantages over other methods.
\end{summary}

@@

\begin{summary}The question of what makes data complex is has been decisively answered. First for IID data in, as maximal Entropy, and then for any data as maximal Kolmogorov Complexity. But there's some dissonance in calling this data information-rich. We are not very interested in watching TV broadcasts of static, or recordings of white noise. This may contain maximal information to a computer, but to us, it's not very interesting. In fact the most interesting data seems to be somewhere in between the zero complexity signal (s tv which is turned off) and the maximum complexity one (analog noise). The question of whether we can capture this notion of `interestingness' in a metric analogous to information entropy and Kolmogorov complexity has occupied research for over half a century, and has produced some surprisingly deep results. This chapter provides a review.
\end{summary}

@@

@@

\begin{summary}Modeling complex objects starts with finding out what the components are. If we know the letters, we can find the words. If we know the words, we can find the sentences. Graphs give us nodes and links for free. How do they combine into functional, recurring units? How can we build up to the global structure?
\end{summary}

@@

\begin{summary}In last chapter we induced motifs for various types of complex networks. We ended by applying the procedure recursively, collapsing upwards across several levels to compress the whole structure.
\end{summary}

@@

\begin{summary}To model graphs in an algorithmic setting we can use any family of computational structures that can turn randomness into graph structures. One recently popular model that has been shown to be effective at producing random examples of any given model is the Deep Belief Network. We investigate their potential to model graph structure, and fit them in to the framework of algorithmic statistics.
\end{summary}

@@

\begin{summary}The classic view of offline machine learning starts with a list of instances from which to learn. These instances can then be represented in some way---preferably a vector of numeric features---and analyzed statistically. In the past decade a situation has emerged and re-emerged that is an ill fit to this paradigm. We are faced with information in a graph: a social network, a web graph, an RDF dataset. We may know our instances, if they correspond to specific nodes in the graph, but the things we know about them are spread throughout the graph. We investigate the problem and provide a bridge to traditional machine learning.
\end{summary}

@@

\begin{summary}Throughout the previous chapters, we have performed many quick tests on small datasets, to illustrate and validate our points. In this chapter we take one use case and investigate more thoroughly, in order to show the full potential of our methods. We choose the domain of citation networks, and investigate various questions.
\end{summary}

@@

Conclusatory chapter.

\appendix
\chapter{Tutorials}
\section{Kolmogorov Complexity}
\section{Minimum Description Length}
\section{Graphs and RDF}
\chapter{Algorithms and techniques}
\chapter{Proofs}

>>>>

\maketitle

\tableofcontents

\chapter{Introduction}
\pb{General explanation of graphs, RDF, Kolmogorov complexity and two-part coding that my mother can understand.}

@@

\section{Complex graphs}

\pb{Complex graphs are not just funny things to look at, they are one of the most robust ways of storing data and communicating concepts. They should/will take a central role in science in the future. Moreso than tabular data.}

@@

\begin{summary} We review the literature on complex graphs, and the approaches that have been used so far to provide statistical analyses on them. We highlight some commond pitfalls in reasoning, and we contrast the status quo with statistical practice on more conventional data.
\end{summary}

\pb{Review paper on complex graph analysis. The current approach to statistics on complex graphs is to propose a an algorithm which generates random graphs, and to show that these graphs have some property in common with the graphs of interest (cf. the Barabasi algorithm and scale freeness). We can do better.}

\chapter{Algorithmic Statistics: Kolmogorov complexity and MDL}

\begin{summary}Minimum Description Length (MDL) and Kolmogorov complexity are two offshoots of the same idea: good compression makes for good statistics. MDL takes this as an axiom and builds on it, while Kolmogorov complexity makes only the assumption that the data is produced by a computational process. The two methods are not incompatible, but they have produced very different results and techniques. In order to have access to the best of both worlds, we start with the rigorous theoretical foundations of Kolmogorov complexity and build a platform that will allow us to fold in the techniques of MDL. We will revisit the principle of two-part coding, considered to be a usable, but often crude, technique in MDL and show that when combined with Kolmogorov complexity it has some deep advantages over other methods.
\end{summary}

\pb{ICALP paper and possible continuation. The end result of this section is that (a) we can use techniques and results from Kolmogorov complexity or MDL in the same context (b) we can use computable approximations and (under resonable assumptions) assume that we are approximating the platonic ideals of Kolmogorov Complexity.}

\section{Algorithmic statistics for graphs}

\pb{Interprets the common practices of analysing complex graphs in terms of algorithmic statistics on a theoretical level. Running example: the Barabasi algorithm and the property of scale freeness.}

@@

\begin{summary}The question of what makes data complex is has been decisively answered. First for IID data---being the datasets with maximal Shannon Entropy---and then for any data as maximal Kolmogorov Complexity. But there's some dissonance in calling this data information-rich. We are not very interested in watching TV broadcasts of static, or recordings of white noise. This may contain maximal information to a computer, but to us, it's not very interesting. In fact the most interesting data seems to be somewhere in between the zero complexity signal (s tv which is turned off) and the maximum complexity one (analog noise). The question of whether we can capture this notion of `interestingness' in a metric analogous to information entropy and Kolmogorov complexity has occupied research for over half a century, and has produced some surprisingly deep results. This chapter provides a review.
\end{summary}

@@

\pb{Review paper on facticity.}

\section{A well-founded, universal approach}

\pb{Our take on facticity.}

@@

\begin{summary} Modeling complex objects starts with finding out what the components are. If we know the letters, we can find the words. If we know the words, we can find the sentences. Graphs give us nodes and links for free. How do they combine into functional, recurring units? How can we build up to the global structure?
\end{summary}

@@

\begin{summary} In the last chapter, we induced motifs for various types of complex networks. We ended by applying the procedure recursively, collapsing upwards across several levels to compress the whole structure into rewrite rules. From this, we can induce recursive rules, and with a rewrite grammar for a given graph.
\end{summary}

@@

\begin{summary}mTo model graphs in an algorithmic setting we can use any family of computational structures that can turn randomness into graphs. One recently popular model that has been shown to be effective at generating random examples is the Deep Belief Network. We investigate their potential to model graph structure, and fit them in to the framework of algorithmic statistics.
\end{summary}

@@

\begin{summary}The classic view of offline machine learning starts with a list of instances from which to learn. These instances can then be represented in some way---preferably a vector of numeric features---and analyzed statistically. In the past decade a situation has emerged and re-emerged that makes for an ill fit to this paradigm. We are faced with information in a graph; perhaps a social network, a web graph or an RDF dataset. We may know our instances, if they correspond to specific nodes in the graph, but the things we know about them are spread throughout the graph. There is no hard line between which parts of our data correspond to which instance. Als data belongs to all instances to verying degrees. We investigate the problem and provide a bridge to traditional machine learning.
\end{summary}

\pb{ESWC paper, of het vervolg daar op. Past niet helemaal in de lijn.}

@@

\begin{summary} Throughout the previous chapters, we have performed many quick tests on small datasets, to illustrate and validate our points. In this chapter we take one use case and investigate more thoroughly, in order to show the full potential of our methods. We choose the domain of citation networks, and investigate various practical questions.
\end{summary}

@@

\pb{Conclusatory chapter. Tells the whole story again, but in terms of the concepts we've introduced. Shows that we did what we set out to do.}

\appendix
\chapter{Tutorials}
\begin{summary}
This appendix provides brief introductions to the subjects that are required to understand this thesis. They illustrate the use of our notation and provide a basis for readers who are well-versed in mathematics and computer science, but lack knowledge of these specific subjects. 
\end{summary}

\section{Kolmogorov complexity}
\section{Minimum description length}
\section{Graphs and RDF-data}
\section{Recommended literature}

\chapter{Symbols and Terms}

\begin{description}
 \item[$K()$]
 \item[$k()$]
 \item[Shannon Entropy]
\end{description}

\chapter{Algorithms and techniques}
\chapter{Proofs}
